### 2020-01-10 paper.tex
pdf diagram of airbnb

% -*- coding: utf-8; -*-
% vim: set fileencoding=utf-8 :
\documentclass[english,submission]{programming}
%% First parameter: the language is 'english'.
%% Second parameter: use 'submission' for initial submission, remove it for camera-ready (see 5.1)

\usepackage{longtable,booktabs}
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\newenvironment{Shaded}{}{}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.49,0.56,0.16}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.53,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.73,0.13,0.13}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.40,0.40,0.40}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.74,0.48,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.73,0.40,0.53}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.10,0.09,0.49}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}

\usepackage[backend=biber]{biblatex}
\addbibresource{references.bib}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\begin{document}

\title{Wildcard: Spreadsheet-Driven Customization of Web Applications}
\subtitle{}% optional

\author{Geoffrey Litt}
\authorinfo{is a PhD student at MIT, researching programming tools for end users and developers.}
\author{Daniel Jackson}
\authorinfo{is a professor in the Department of Electrical Engineering and Computer Science at MIT, associate director of CSAIL, and a MacVicar Fellow. He leads the Software Design Group.}
\affiliation{Massachusetts Institute of Technology}

\keywords{end-user programming, software customization, web browser extensions} % please provide 1--5 keywords

%%%%%%%%%%%%%%%%%%
%% These data MUST be filled for your submission. (see 5.3)
\paperdetails{
  %% perspective options are: art, sciencetheoretical, scienceempirical, engineering.
  %% Choose exactly the one that best describes this work. (see 2.1)
  perspective=art,
  %% State one or more areas, separated by a comma. (see 2.2)
  %% Please see list of areas in http://programming-journal.org/cfp/
  %% The list is open-ended, so use other areas if yours is/are not listed.
  area={Programming environments, Visual and live programming},
  %% You may choose the license for your paper (see 3.)
  %% License options include: cc-by (default), cc-by-nc
  % license=cc-by-sa,
}
%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%
%% These data are provided by the editors. May be left out on submission.
%\paperdetails{
%  submitted=2016-08-10,
%  published=2016-10-11,
%  year=2016,
%  volume=1,
%  issue=1,
%  articlenumber=1,
%}
%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Please go to https://dl.acm.org/ccs/ccs.cfm and generate your Classification
% System [view CCS TeX Code] stanz and copy _all of it_ to this place.
%% From HERE
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10002944.10011122.10003459</concept_id>
<concept_desc>General and reference~Computing standards, RFCs and guidelines</concept_desc>
<concept_significance>300</concept_significance>
</concept>
<concept>
<concept_id>10010405.10010476.10010477</concept_id>
<concept_desc>Applied computing~Publishing</concept_desc>
<concept_significance>300</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[300]{General and reference~Computing standards, RFCs and guidelines}
\ccsdesc[500]{Applied computing~Publishing}

% To HERE
%%%%%%%%%%%%%%%%%%%%%%%

\maketitle

% Please always include the abstract.
% The abstract MUST be written according to the directives stated in 
% http://programming-journal.org/submission/
% Failure to adhere to the abstract directives may result in the paper
% being returned to the authors.
\begin{abstract}
Many Web applications do not meet the particular needs of their users.
Browser extensions and user scripts offer a way to customize web
applications, but most people do not have the programming skills to
implement their own extensions.

We present the idea of \emph{spreadsheet-driven customization}: enabling
end users to customize existing applications using a live spreadsheet
view of the data inside the application. By manipulating the
spreadsheet, users can implement a wide variety of customizations,
ranging from sorting lists of search results to displaying related data
from other web services, without doing any traditional programming.

We built a prototype system called Wildcard that implements
spreadsheet-driven customization as a web browser extension. Through
concrete examples, we demonstrate that Wildcard has both a low barrier
to entry for beginners and enough flexibility to solve many useful
problems. We also show that Wildcard can work with real existing
websites, by extracting structured data using web scraping techniques.

For an online version of this paper with videos demonstrating the interface, see \url{https://www.geoffreylitt.com/wildcard/}
\end{abstract}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

In 2012, the travel site Airbnb removed the ability to sort listings by
price. Users could still filter by price range, but could no longer view
the cheapest listings first. Many users complained on online message
boards that the change seemed hostile to users. ``It's so
frustrating!..What is the logic behind not having this function?'' said
one user on the
\href{https://community.withairbnb.com/t5/Hosting/Sorting-listing-by-price/td-p/559404}{Airbnb
support forum}. Alas, the feature remains missing to this day.

This is a familiar situation in a world of web applications that are
frequently updated without user consent. Sometimes there is a browser
extension or user script that fixes an issue, and if the user is both
motivated and skilled they might even be able to implement their own
fix. But for most people, the only recourse is to complain to the
developers and pray that someone listens---or more likely, to simply
give up. While many have become used to this status quo, we see it as a
tremendous waste of the openness of the Web platform. Back in 1977, in
\emph{Personal Dynamic Media} \autocite{kay1977}, Alan Kay originally
envisioned personal computing as a medium that let a user ``mold and
channel its power to his own needs,'' but today software feels more like
concrete than clay.

In this paper, we present \emph{spreadsheet-driven customization}, a
technique for making software customization more accessible to end
users. The core idea is to show the data inside the application in a
spreadsheet which maintains a live connection to the original UI. When
the user manipulates the spreadsheet, the UI is instantly modified, and
vice versa. We built a research prototype called Wildcard that
implements spreadsheet-driven customization as a web browser extension,
showing a spreadsheet panel in the context of existing sites.

Spreadsheets have proven to be a widely successful computing platform
for non-programmers. They provide a low barrier to entry for beginners,
while allowing experts to perform complex computations. The 2D grid is
simple to grasp, but general enough to represent many types of data.
Prior work \autocite{mccutchen2016,benson2014,chang2014} has used these
benefits to enable end users to easily create web applications that use
spreadsheets as a backing data layer.

Spreadsheet-driven customization applies these exact same benefits in a
different context: customizing existing software, rather than building
new software from scratch. We inherit the low floor and high ceiling of
spreadsheets: small tweaks like sorting a list of data in an app can be
done with a single click, while more complex customizations, like
joining in related data from a web API, are supported by a rich formula
system. The tabular data format is versatile, supporting data from a
wide range of applications. Spreadsheet-driven customization does not
require that the application is actually backed by a spreadsheet---it
merely presents a model of the internal state of the application, which
can be exposed by the application itself or derived from the user
interface by third parties. This means that spreadsheet-driven
customization works with existing applications that people already use.
It also allows for exposing a common data abstraction across
applications, enabling users to have a consistent mechanism for
manipulating data from all of their applications.

Prior work \autocite{huynh2006} has shown that the process of extracting
structured data from existing user interfaces can be confusing and
unfamiliar for end users. In order to make the customization experience
as straightforward as possible, we envision a tiered architecture that
hides this complexity from end users. Wildcard provides a mechanism for
programmers to write \emph{site adapters}, which use web scraping
techniques to extract structured data from the DOM and AJAX requests of
existing applications. End users only interact with the structured
spreadsheet, providing a predictable experience with clear affordances
for which customizations are possible.

Wildcard is currently an early prototype, with incomplete features and
limited coverage of sites. We plan to continue building site adapters
and testing the system with our own use cases, to better understand how
well the spreadsheet abstraction maps to real websites and customization
needs. Eventually we also plan to release the tool publicly, to
determine how end users choose to use spreadsheet-driven customization,
what usability challenges emerge, and how feasible it is for programmers
to build and maintain site adapters.

\hypertarget{demo-booking-a-trip-with-wildcard}{%
\section{Demo: booking a trip with
Wildcard}\label{demo-booking-a-trip-with-wildcard}}

To get a sense of how it might feel to use Wildcard, let's see an
example of someone using it to help with booking a trip on the travel
sites Airbnb and Expedia. These demos are best viewed in the online
version of this paper, which includes videos:
\url{https://geoffreylitt.com/wildcard}

\begin{figure}
\hypertarget{fig:airbnb-demo}{%
\centering
\includegraphics{media/airbnb-demo-300dpi.png}
\caption{Using Wildcard to augment the Airbnb page for searching travel
accommodations}\label{fig:airbnb-demo}
}
\end{figure}

The user starts by opening up the Airbnb search listings page to look
for a place to stay. The page looks nice and mostly works well, but is
missing some key features. As mentioned before, this page doesn't allow
the user to sort by price. It also doesn't let them filter by user
rating. Using Wildcard, the user can add these small features, while
leaving the page's design and the rest of its functionality unchanged
(shown in fig.~\ref{fig:sorting}).

After manipulating the data, the user can close the table view and
continue using the website with its original visual design. The table
view offers a way to change the data backing a page, but does not need
to replace the original interface entirely.

Most websites that show lists of data also offer actions that can be
taken on a row in the table, like adding an item to a shopping cart.
Wildcard has the ability to make these ``row actions'' available in the
data table if the site adapter implements them. The main advantage this
provides is the ability to easily perform a row action in bulk across
multiple rows.

For example, it's tedious on Airbnb to click on listings one by one to
add them to a list of favorites. Using Wildcard, the user can just
select multiple rows and favorite all of them with one click. Similarly,
we can also open the detailed pages for many listings in new tabs.
(Shown in fig.~\ref{fig:actions})

\begin{figure}
\hypertarget{fig:expedia-demo}{%
\centering
\includegraphics{media/expedia-demo-300dpi.png}
\caption{Using Wildcard to augment the Expedia page for booking a
flight}\label{fig:expedia-demo}
}
\end{figure}

Here we've presented just a few possibilities for how to use Wildcard.
We think the interactive data table offers a flexible computational
model that can support a wide range of other useful modifications, all
while remaining familiar and easy to use.

\hypertarget{system-architecture}{%
\section{System Architecture}\label{system-architecture}}

Wildcard is written in Typescript. It is currently injected into pages
using the \href{https://www.tampermonkey.net/}{Tampermonkey} userscript
manager, but in the future we plan to deploy it as a standalone browser
extension to make it easier to install.

In order to maximize extensibility, Wildcard is implemented as a small
core program along with several types of plugins: site adapters,
formulas, and cell renderers/editors. The core contains functionality
for displaying the data table and handling user interactions, and the
table implementation is built using the
\href{https://handsontable.com/}{Handsontable} Javascript library.

\begin{figure}
\hypertarget{fig:architecture}{%
\centering
\includegraphics{media/architecture-clean.png}
\caption{The architecture of the Wildcard
system}\label{fig:architecture}
}
\end{figure}

Site adapters specify the bidirectional connection between the web page
and its structured data representation.

For extracting data from the page and getting it into structured form,
Wildcard provides ways to concisely express web scraping logic. For
example, here is a code snippet for extracting the name of an Airbnb
listing:

\begin{Shaded}
\begin{Highlighting}[]
  \OperatorTok{\{}
\NormalTok{    fieldName}\OperatorTok{:} \StringTok{"name"}\OperatorTok{,} \CommentTok{// The name of the data field}
\NormalTok{    readOnly}\OperatorTok{:} \KeywordTok{true}\OperatorTok{,}    \CommentTok{// Whether the user can edit the field}
\NormalTok{    type}\OperatorTok{:} \StringTok{"text"}\OperatorTok{,}      \CommentTok{// The type of the field}
    \CommentTok{// Given the DOM element for the entire listing,}
    \CommentTok{// return the DOM element representing this field}
\NormalTok{    el}\OperatorTok{:}\NormalTok{ (row) }\KeywordTok{=>} \VariableTok{row}\OperatorTok{.}\FunctionTok{querySelector}\NormalTok{(}\VerbatimStringTok{\textasciigrave{}.}\SpecialCharTok{$\{}\NormalTok{titleClass}\SpecialCharTok{\}}\VerbatimStringTok{\textasciigrave{}}\NormalTok{)}\OperatorTok{,}
  \OperatorTok{\}}
\end{Highlighting}
\end{Shaded}

Sometimes important data is not shown in the UI, making it impossible to
scrape from the DOM. To address this, we have also prototyped mechanisms
for site adapters to observe AJAX requests made by the browser and
extract data directly from JSON responses. This mechanism was used to
implement the Airbnb Walkscore example, since latitude and longitude
aren't shown in the Airbnb UI, but they are available in AJAX responses.
We also plan to add the ability for site adapters to scrape data across
multiple pages in paginated lists of results (as explored in prior work
\autocite{huynh2006}).

The site adapter also needs to support the reverse direction: sending
updates from the table to the original page. Most DOM manipulation is
not performed directly by the site adapter; instead, the adapter
specifies how to find the divs representing data rows, and the core
platform mutates the DOM to reflect the table state. The only exception
is row actions (like favoriting an Airbnb listing), which are
implemented as imperative Javascript functions that can can mutate the
DOM, simulate clicks on buttons, etc.

\hypertarget{design-principles}{%
\section{Design principles}\label{design-principles}}

The idea of spreadsheet-driven customization is guided by several design
principles, inspired by prior work and our own experimentation. We think
these principles can also broadly inform the design of end user
programming tools, especially those that enable users to customize
existing software.

\hypertarget{expose-a-universal-data-structure}{%
\subsection{Expose a universal data
structure}\label{expose-a-universal-data-structure}}

Today, most personal computing consists of using applications, which
bundle together behavior and data to provide some set of functionality.
While there are limited points of interoperability, applications
generally are designed to operate independently of one another. Once a
user has gotten past learning the basic idioms of modern computing like
windows and scrollbars, most of the effort invested in learning to use
one application does not carry over to using other applications.

Computing does not need to be organized this way. For example, UNIX
offers a compelling alternative design: many small single-purpose
utilities, all of which manipulate a universal format of text streams.
The universal format creates a high degree of leverage from tools: users
can get a lot of utility from deeply mastering a text editor and some
text manipulation utilities, because these tools can be applied to
nearly any task. As just one example, a user's preferred text editor can
even serve as an interactive input mechanism in shell programs, e.g.~for
editing git commit messages.

Spreadsheet-driven customization aims to port this UNIX philosophy to
the world of isolated Web applications, by creating a consistent data
structure to represent the data inside many applications. In UNIX, the
universal format is a text stream; in Wildcard, it is a relational
table: a simple abstraction generic enough to describe the data used in
many different applications. Because Wildcard maps the data from all
applications to the table format, users can invest in mastering the
Wildcard table editor, the formula language, and cell editor UIs, and
reuse those same tools to customize many different applications.

This idea relates to Beaudouin-Lafon and Mackay's notion of
\emph{polymorphic interface instruments} \autocite{beaudouin-lafon2000}:
UI elements that can be used in different contexts, like a color picker
that can be used in many different drawing applications. diSessa has
also noted the connection between literacy and the genericness of a
medium. Textual literacy rests on the fact that writing can be adapted
to many different genres and uses \autocite{disessa2000}; if people
needed to relearn reading and writing from scratch when switching from
essays to emails, the medium would lose most of its potency. We think
providing generic tools is especially important for software
customization, because the most common barrier to customizing software
is not having enough time \autocite{mackay1991}---it's more likely that
people will customize software regularly if they can reuse the same tool
across many applications.

This design principle leads to several challenges. First, any universal
abstraction has its constraints, and can't necessarily naturally express
the data in every application. We plan to explore the limits of the
table abstraction further, by trying to build adapters for more sites
with varied data formats. We expect that many types of data can fit
fairly naturally into tables: lists of search results, news articles,
and messages can all be seen as relations. On sites that use document
structures (e.g.~Google Docs) or graph structures (e.g.~social friend
graphs), it may prove more challenging to map internal data to this
abstraction.

Another challenge is ensuring a clear mapping in the user's mind between
the spreadsheet and the original page. Wildcard provides live visual
cues as the user navigates the data table (similar to the highlighting
provided by DOM inspectors in browser developer tools). In our own
usage, we have found that this live highlighting makes it very clear how
the two representations map to each other.

\hypertarget{low-floor-high-ceiling}{%
\subsection{Low floor, high ceiling}\label{low-floor-high-ceiling}}

Seymour Papert advocated for programming systems with a ``low floor,''
making it easy for novices to get started, as well as a ``high
ceiling,'' providing a large range of possibilities for more
sophisticated users \autocite{resnick2016}. Our goal is for
spreadsheet-driven customization to meet both of these criteria.
Although we need more testing with real users to know for sure, we think
Wildcard clearly provides a low floor, and likely has a flexible enough
foundation to eventually provide a high ceiling.

One of the most interesting properties of spreadsheets is the amount of
value they can provide to users who are aware of only a tiny sliver of
their functionality. Not only can a novice perform some basic activities
in a spreadsheet with almost no training (e.g., storing tables of
numbers or computing simple sums), these activities are often actually
valuable for the user! The fact that useful tasks can be performed early
on supports the user's natural motivation to continue using the tool,
and to eventually learn its more powerful features if needed
\autocite{nardi1991}. In contrast, many traditional programming systems
require an enormous upfront investment of time and practice before
someone is able to write a program that actually helps them achieve a
real task.

As part of ensuring a low floor for spreadsheet-driven customization, we
have focused on including genuinely valuable features for novices. For
example, a user can sort a table with a single click, or simply type in
some annotations. We would expect many Wildcard users to start out using
these simpler features before potentially moving on to more
sophisticated features like formulas.

Another aspect of providing a low floor is providing an ``in-place
toolchain'' \autocite{inkandswitch2019}---minimizing the effort of
moving from using to customizing, by making customization tools
available in the same environment where the user is already using the
software. This quality is distinct from the level of technical skill
needed to use the tool: for example, setting up a workflow trigger in an
end user programming system like \href{https://ifttt.com/}{IFTTT} does
not require much technical skill, but does require leaving the user's
normal software and entering a separate environment; conversely, running
a Javascript snippet in the browser console requires programming skills,
but can be done immediately and casually in the flow of using a website.

\begin{longtable}[]{@{}llc@{}}
\toprule
\endhead
& \emph{In-place} & \emph{Not in-place}\tabularnewline
\emph{End user friendly} & \textbf{Wildcard} & IFTTT\tabularnewline
\emph{Requires programming} & browser JS console & forking open source
software\tabularnewline
\bottomrule
\end{longtable}

Wildcard provides an in-place toolchain---on any site that supports
Wildcard, the user is one click away from starting to tweak the site
with minimal friction. There is no need for them to switch into another
environment or figure out how to extract data from the page. Once the
user starts editing, Wildcard also provides live feedback to help users
understand the changes they are making. Even if a user isn't yet totally
familiar with Wildcard, they can learn to use the system by trying
things out and seeing what happens.

Since we have still only built several site adapters and demos, it's
still too early to tell how high the ceiling is for the customizations
that can be achieved with Wildcard, but we think that with enough
operators the formula language could eventually support a wide variety
of customizations. We plan to explore this aspect further by trying to
solve more real problems with the system and observing where limitations
emerge in practice.

\hypertarget{build-for-multiple-tiers-of-users}{%
\subsection{Build for multiple tiers of
users}\label{build-for-multiple-tiers-of-users}}

Real-world spreadsheet usage in offices is highly collaborative: most
users just perform simple changes, while a few coworkers help with
writing more complex formulas or even programming macros
\autocite{nardi1990}. Inspired by this, we aim to make
spreadsheet-driven customization a collaborative activity that combines
the different abilities of many users in a collaborative ecosystem.

The main way we do this is by separating website customization into two
separate stages: structured data extraction, and using the resulting
spreadsheet. The first stage is currently only available to programmers
who can code site adapters in Javascript, whereas the second stage is
available for any non-programmer end user. This architecture frees end
users from needing to think about data extraction, and enables a
community of end users to reuse the efforts of programmers building site
adapters.

The group of users building adapters does not necessarily need to be
limited only to programmers. In the future we might explore enabling end
users to also create site adapters, drawing on related work on enabling
end users to extract structured data from websites
\autocite{chasins2018,huynh2006}. But even in that case, there would
still be a separation between motivated, tech-savvy end users building
adapters and more casual end users just using the spreadsheet view.

Another stakeholder to consider is the first party developers of the
original software. Spreadsheet-driven customization does not depend on
cooperation from first-party website developers: a third party
programmer can write an adapter for any website, which can access any
information available in the browser. On the other hand, if first
parties were to expose structured data from their applications, it would
avoid the need for adapters and generally make customization a lot
easier. We think there are compelling reasons for first parties to
consider doing this. Providing Wildcard support would allow users to
build extensions to fulfill their own feature requests. It also would
not necessarily require much effort: adding Wildcard support would be
more straightforward for a first-party than a third-party because they
have direct access to the structured data in the page. There is also
precedent for first parties implementing an official client extension
API in response to user demand: for several years, Google maintained an
official extension API in Gmail for Greasemonkey scripts to
use.\footnote{Incidentally, since then, third parties have continued to
  maintain stable Gmail extension APIs used by many browser extensions
  \autocite{streak,talwar2019}, illustrating the potential of
  collaboratively maintaining third party adapters.}

\hypertarget{related-work}{%
\section{Related work}\label{related-work}}

\hypertarget{malleable-software}{%
\subsection{Malleable software}\label{malleable-software}}

In the broadest sense, Wildcard is inspired by systems aiming to make
software into a dynamic medium where end users frequently create and
modify software to meet their own needs, rather than only consuming
applications built by programmers. These systems include Smalltalk
\autocite{kay1977}, Hypercard \autocite{hypercard2019} , Boxer
\autocite{disessa1986}, Webstrates \autocite{klokmose2015}, and
Dynamicland \autocite{victor}.\footnote{The project's name Wildcard
  comes from the internal pre-release name for Hypercard, which doubly
  inspired our work by promoting both software modification by end users
  and the ideas behind the Web.}

While similar in high-level goals, Wildcard employs a different solution
strategy from these projects: whereas they generally require building
new software from scratch for that environment, Wildcard instead aims to
maximize the malleability of already existing software. This approach
has the pragmatic benefit of being immediately useful in more
situations, although it also requires working within more rigid
constraints.

With substantial future work, Wildcard could become more similar to
these other projects, growing from a platform for tweaking existing
software into a platform for building new software from scratch. This
would likely end up resembling existing tools for building
spreadsheet-driven applications (discussed more below), but with an
extra focus on customizability by end users of the software.

\hypertarget{web-customization}{%
\subsection{Web customization}\label{web-customization}}

Wildcard's goals are closely shared with other systems that provide
interfaces in the browser for end users to augment and customize
websites while using them.

\hypertarget{structured-augmentation}{%
\subsubsection{Structured augmentation}\label{structured-augmentation}}

Wildcard's approach is most similar to a class of tools that identify
structured data in a web page, and use that structure to support end
user modification of the page.

Sifter \autocite{huynh2006} enables users to sort and filter lists of
data on web pages, providing a result similar to Wildcard's sort and
filter functionality. The underlying mechanism is also similar: Sifter
extracts structured data from the page to enable its user-facing
functionality. Wildcard aims to extend this approach to support much
broader functionality than just sorting and filtering. In support of
this goal, Wildcard also shows the structured data table directly to the
user, whereas Sifter only shows sort and filter controls, without
revealing the underlying data table. The extraction mechanism is also
different: Sifter uses a combination of automated heuristics and
interactive user feedback to extract data, whereas Wildcard currently
relies on programmers creating wrappers for extracting structured data,
likely leading to higher quality extraction but on fewer sites.

Thresher \autocite{hogue2005} enables users to create wrappers which map
unstructured website content to Semantic Web content. Like Wildcard and
Sifter, Thresher augments the experience of original page based on
identifying structure: once semantic content has been identified, it
creates context menus in the original website which allow users to take
actions based on that content. Wildcard and Thresher share an overall
approach but focus on complementary parts of the problem: Thresher aims
to enable end users to create content wrappers, but the actions
available on the structured data are created by programmers; conversely,
Wildcard delegates wrapper creation to programmers but gives end users
more flexibility to use the structured data in an open-ended way.

\hypertarget{sloppy-augmentation}{%
\subsubsection{Sloppy augmentation}\label{sloppy-augmentation}}

``Sloppy programming'' \autocite{little2010} tools like Chickenfoot
\autocite{bolin2005} and Coscripter \autocite{leshed2008} enable users
to create scripts that perform actions like filling in text boxes and
clicking buttons, without directly interacting with the DOM. Users
express the desired page elements in natural, informal terms
(e.g.~writing ``the username box'' to represent the textbook closest to
the label ``username''), and then using heuristics to determine which
elements most likely match the user's intent. This approach allows for
expressing a wide variety of commands with minimal training, but it also
has downsides. It is difficult to know whether a command will
consistently work over time (in addition to changes to the website,
changes to the heuristics can also cause problems), and it is not easy
for users to discover the space of possible commands
\autocite{little2010}.

Wildcard offers a sharp contrast to sloppy programming, instead choosing
to expose a high degree of structure through the familiar spreadsheet
table. Wildcard offers more consistency: for example, clicking a sort
header will always work correctly as long as the site adapter is
maintained. Wildcard also offers clearer affordances for what types of
actions are possible, or, crucially, what actions are \emph{not}
possible, which is useful to know. On the other hand, Wildcard cannot
offer coverage of all websites, and has a narrower range of possible
actions than sloppy tools. We expect that with enough site adapters and
formulas, these downsides can be mitigated.

\hypertarget{spreadsheet-based-app-builders}{%
\subsection{Spreadsheet-based app
builders}\label{spreadsheet-based-app-builders}}

Prior work has made the powerful realization that a spreadsheet can
serve as an end-user-friendly backing data store and computation layer
for an interactive web application. Research projects like Object
Spreadsheets \autocite{mccutchen2016}, Quilt \autocite{benson2014},
Gneiss \autocite{chang2014}, and Marmite \autocite{wong2007}, as well as
commercial tools like Airtable Blocks \autocite{zotero-79} and Glide
\autocite{zotero-81} allow users to view data in a spreadsheet table,
compute over the data using formulas, and connect the table to a GUI.
Because many users are already familiar with using spreadsheets, this
way of creating applications tends to be far easier than traditional
software methods; for example, in a user study of Quilt, many people
were able to create applications in under 10 minutes, even if they
expected it would take them many hours.

Wildcard builds on this idea, but applies it to modifying existing
applications, rather than building new applications from scratch. For
many people, we suspect that tweaking existing applications provides
more motivation as a starting point for programming than creating a new
application from scratch.

An important design decision for tools in this space is how to deviate
from traditional spreadsheets like Microsoft Excel or Google Sheets.
Quilt and Glide use existing spreadsheet software as a backend,
providing maximum familiarity for users, and even compatibility with
existing spreadsheets. Gneiss has its own spreadsheet implementation
with additional features useful for building GUIs. Marmite provides a
live data view that resembles a spreadsheet, but programming is actually
done using a separate data flow pane rather than spreadsheet formulas.
(Marmite's approach led to some confusion in a user study, because users
expected behavior more similar to spreadsheets \autocite{wong2007}.)
Airtable deviates the furthest: although the user interface resembles a
spreadsheet, the underlying structure is a relational database with
typed columns. Wildcard's table is most similar to Airtable; the
structure of a relational table is most appropriate for most data in
websites, and we have not yet found a need for arbitrary untyped cells.

\hypertarget{web-scraping-data-extraction}{%
\subsection{Web scraping / data
extraction}\label{web-scraping-data-extraction}}

Web scraping tools focus on extracting structured data out of
unstructured web pages. Web scraping is closely related to the
implementation of Wildcard, but has different end goals: web scraping
generally extracts static data for processing in another environment,
whereas Wildcard modifies the original page by maintaining a
bidirectional connection between the extracted data and the page.

Web scraping tools differ in how much structure they attempt to map onto
the data. Some tools like Rousillon \autocite{chasins2018} extract data
in a minimally structured relational format; other tools like Piggy Bank
\autocite{huynh2005} more ambitiously map the data to a rich semantic
schema. In Wildcard, we chose to avoid schemas, in order to minimize the
work associated with creating a site adapter.

In the future, we might be able to integrate web scraping tools to help
create more reliable site adapters for Wildcard with less work, and to
open up adapter creation to end users. Sifter was built on top of the
Piggy Bank scraping library, suggesting precisely this type of
architecture where web scraping tools are used to support interactive
page modification.

\hypertarget{future-work}{%
\section{Future work}\label{future-work}}

There are still many open questions about spreadsheet-driven
customization, which we hope to answer through targeted development and
usage of the Wildcard prototype.

The most important question is whether the computational model can
support a large variety of useful customizations. While initial demos
are promising, we need to develop more site adapters and use cases to
explore this further. We plan to continue privately testing the system
with our own needs, and to eventually deploy the tool publicly, once the
API is stable enough and can support a critical mass of sites and use
cases. We also plan to run usability studies to evaluate and improve the
design of the tool.

We suspect that two areas of the current model may prove too limiting.
First, Wildcard's data model only shows a single table at a time,
without any notion of relationships between tables. A richer data model
with foreign keys might be necessary to support certain use cases. For
designing a tabular interface on top of a richer data model, we could
learn from the interface of Airtable which shows related objects in a
table cell, or add nested rows as used in other systems
\autocite{mccutchen2016,bakke2016}.

Second, there is no mechanism for end users to express imperative
workflows (e.g.~``click this button, and then\ldots{}''); they can only
write formulas that return data and then inject the resulting data into
the page. If it proves useful, we might consider adding a system for
expressing workflows like this, although it would substantially alter
the computational model and it's not clear how it would fit together
with the existing table view.

Another open question is how efficiently site adapters can be created
for real websites. By making adapters for sites with different types of
data in different domains, we plan to explore how automated heuristics
can speed up the adapter creation process, and to create abstractions
that make it easier for programmers to efficiently create new robust
adapters.

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

In this paper, we have presented \emph{spreadsheet-driven
customization}, a way to make software customization more accessible to
end users by showing the data inside the application in a spreadsheet.
We hope that this technique contributes to making the Web into a more
dynamic medium that users can mold to their own needs.

We plan to continue developing the Wildcard prototype and to eventually
deploy it as an open-source tool. To receive future updates on Wildcard
and notifications about a public release, \href{}{sign up for the email
newsletter}.

We are also looking for private beta testers. If you have an idea for
how you might want to use Wildcard, please
\href{mailto:glitt@mit.edu}{get in touch}. We would love to hear about
your needs and help find ways to use Wildcard to solve them.

\acks
\printbibliography

\end{document}

% Local Variables:
% TeX-engine: luatex
% End: