### 2020-01-10 paper.tex
done for the day

% -*- coding: utf-8; -*-
% vim: set fileencoding=utf-8 :
\documentclass[english,submission]{programming}
%% First parameter: the language is 'english'.
%% Second parameter: use 'submission' for initial submission, remove it for camera-ready (see 5.1)

\usepackage{longtable,booktabs}
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\newenvironment{Shaded}{}{}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.49,0.56,0.16}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.53,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.73,0.13,0.13}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.40,0.40,0.40}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.74,0.48,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.73,0.40,0.53}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.10,0.09,0.49}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}

\usepackage[backend=biber]{biblatex}
\addbibresource{references.bib}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\begin{document}

\title{Wildcard: Spreadsheet-Driven Customization of Web Applications}
\subtitle{}% optional

\author{Geoffrey Litt}
\authorinfo{is a PhD student at MIT, researching programming tools for end users and developers.}
\author{Daniel Jackson}
\authorinfo{is a professor in the Department of Electrical Engineering and Computer Science at MIT, associate director of CSAIL, and a MacVicar Fellow. He leads the Software Design Group.}
\affiliation{Massachusetts Institute of Technology}

\keywords{end-user programming, software customization, web browser extensions} % please provide 1--5 keywords

%%%%%%%%%%%%%%%%%%
%% These data MUST be filled for your submission. (see 5.3)
\paperdetails{
  %% perspective options are: art, sciencetheoretical, scienceempirical, engineering.
  %% Choose exactly the one that best describes this work. (see 2.1)
  perspective=art,
  %% State one or more areas, separated by a comma. (see 2.2)
  %% Please see list of areas in http://programming-journal.org/cfp/
  %% The list is open-ended, so use other areas if yours is/are not listed.
  area={Programming environments, Visual and live programming},
  %% You may choose the license for your paper (see 3.)
  %% License options include: cc-by (default), cc-by-nc
  % license=cc-by-sa,
}
%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%
%% These data are provided by the editors. May be left out on submission.
%\paperdetails{
%  submitted=2016-08-10,
%  published=2016-10-11,
%  year=2016,
%  volume=1,
%  issue=1,
%  articlenumber=1,
%}
%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Please go to https://dl.acm.org/ccs/ccs.cfm and generate your Classification
% System [view CCS TeX Code] stanz and copy _all of it_ to this place.
%% From HERE
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10011007.10011006.10011066.10011069</concept_id>
<concept_desc>Software and its engineering~Integrated and visual development environments</concept_desc>
<concept_significance>500</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Software and its engineering~Integrated and visual development environments}
% To HERE
%%%%%%%%%%%%%%%%%%%%%%%

\maketitle

% Please always include the abstract.
% The abstract MUST be written according to the directives stated in 
% http://programming-journal.org/submission/
% Failure to adhere to the abstract directives may result in the paper
% being returned to the authors.
\begin{abstract}
Many Web applications do not meet the particular needs of their users.
Browser extensions and user scripts offer a way to customize web
applications, but most people do not have the programming skills to
implement their own extensions.

We present the idea of \emph{spreadsheet-driven customization}: enabling
end users to customize existing applications using a live spreadsheet
view of the data inside the application. By manipulating the
spreadsheet, users can implement a wide variety of customizations,
ranging from sorting lists of search results to displaying related data
from other web services, without doing any traditional programming.

We built a prototype system called Wildcard that implements
spreadsheet-driven customization as a web browser extension. Through
concrete examples, we demonstrate that Wildcard has both a low barrier
to entry for beginners and enough flexibility to solve many useful
problems. We also show that Wildcard can work with real existing
websites, by extracting structured data using web scraping techniques.

This paper is best viewed in the online version, which includes videos demonstrating the interface: \url{https://www.geoffreylitt.com/wildcard/}
\end{abstract}

\hypertarget{sec:introduction}{%
\section{Introduction}\label{sec:introduction}}

Web applications often don't match the particular needs of their users.
Sometimes there is a browser extension available to patch an issue, and
if the user is a programmer they might be able to fix it themselves. But
for most people, the only recourse is to complain to the developers, or
more likely, to simply give up. Back in 1977, in \emph{Personal Dynamic
Media} \autocite{kay1977}, Alan Kay envisioned personal computing as a
medium that let a user ``mold and channel its power to his own needs,''
but today, software behaves more like concrete than clay.

In this paper, we present \emph{spreadsheet-driven customization}, a
technique that enables end users to customize software without doing any
traditional programming. The idea is to augment an application's UI with
a spreadsheet that is synchronized with the application's data. When the
user manipulates the spreadsheet, the underlying data is modified and
the changes are propagated to the UI, and vice versa. We have
implemented this technique in a prototype browser extension called
Wildcard. Section~\ref{sec:demos} presents demos of using Wildcard to
augment real websites in useful ways, and
Section~\ref{sec:implementation} describes the implementation of the
extension.

\begin{figure}
\hypertarget{fig:overview}{%
\centering
\includegraphics{media/overview-300dpi.png}
\caption{An overview of spreadsheet-driven
customization}\label{fig:overview}
}
\end{figure}

Prior work \autocite{mccutchen2016,benson2014,chang2014} has enabled end
users to create ``spreadsheet-driven applications'' which use
spreadsheets as a backing data layer. Spreadsheet-driven
\emph{customization} applies this idea in a different context:
customizing existing software, rather than building new software from
scratch. Our technique does not require that the application actually be
backed by a spreadsheet; it merely uses the spreadsheet as an interface
for viewing and modifying the internal state of the application.

Spreadsheet-driven customization provides an easy entry point for end
users, since small tweaks like sorting data can be performed with a
single click. At the same time, it also supports a variety of richer
customizations, like adding private annotations to a webpage or joining
in related data from a web API. In Section~\ref{sec:design-principles},
we elaborate on this breadth of use cases, as well as other design
principles guiding our work.

This approach requires extracting structured data from the user
interfaces of existing applications, but we hide the complexity of data
extraction from end users. Programmers write \emph{site adapters} which
use web scraping techniques to extract structured data from existing
applications and map them to the spreadsheet table. End users only
interact with the structured spreadsheet, providing a straightforward
customization experience. In Section~\ref{sec:related-work}, we explain
this architecture in more depth, and describe how Wildcard relates to
existing work on malleable software, web customization,
spreadsheet-driven applications, and web scraping.

The Wildcard extension is currently an early research prototype. We plan
to continue testing the system with our own use cases to explore how the
spreadsheet abstraction maps to real websites and customization needs.
Eventually we plan to release the tool publicly, in order to study real
applications, discover usability challenges, and to test the feasibility
of programmers building and maintaining site adapters.

\hypertarget{sec:demos}{%
\section{Demo: booking a trip with Wildcard}\label{sec:demos}}

To demonstrate the user experience of Wildcard, here is an example of
using it to help with booking a trip.{ These demos are best viewed as
videos in the online version of this paper
(\url{https://www.geoffreylitt.com/wildcard}).}

\begin{figure}
\hypertarget{fig:airbnb-demo}{%
\centering
\includegraphics{media/airbnb-demo-300dpi.png}
\caption{Using Wildcard to augment the Airbnb search page for booking
accommodations}\label{fig:airbnb-demo}
}
\end{figure}

\hypertarget{customizing-search-results}{%
\subsection{Customizing search
results}\label{customizing-search-results}}

In 2012, the travel site Airbnb removed the ability to sort
accommodation searches by price. Users could still filter by price
range, but could no longer view the cheapest listings first. Many users
complained that the change seemed hostile to users. ``It's so
frustrating!..What is the logic behind not having this function?'' said
one user on the
\href{https://community.withairbnb.com/t5/Hosting/Sorting-listing-by-price/td-p/559404}{Airbnb
support forum}. Alas, the feature remains missing to this day.

Using Wildcard, the user can fix this omission, while leaving the page's
design and the rest of its functionality unchanged.{
Figure~\ref{fig:airbnb-demo} shows an overview of augmenting the Airbnb
site.} First, the user opens up the Wildcard panel, which shows a table
corresponding to the search results in the page. As they click around in
the table, the corresponding row in the page is highlighted so they can
see the connection between the views.

Then, the user clicks on the price column header to sort the spreadsheet
and the Airbnb UI by price{ (Figure~\ref{fig:airbnb-demo}B)}. They also
filter to listings with a user rating above 4.5 (another feature missing
in the original Airbnb UI).

After manipulating the data, the user closes the table view and
continues using the website with its original visual design. The table
view offers a way to change the data backing a page, but does not need
to replace the original interface entirely.

Most websites that show lists of data also offer actions on rows in the
table, like adding an item to a shopping cart. Wildcard has the ability
to make these ``row actions'' available in the data table through the
site adapter. In the Airbnb UI, saving multiple listings to a Favorites
list requires tediously clicking through them one by one. Using Wildcard
row actions, the user can select multiple rows and favorite all of them
with a single click{ (Figure~\ref{fig:airbnb-demo}D)}. Similarly, the
user can open the detailed pages for many listings at once.

Next, the user wants to jot down some notes about each listing. To do
this, they type some notes into an additional column in each listing
row, and the notes appear inside the listings in the original UI{
(Figure~\ref{fig:airbnb-demo}A)}. The annotations are saved in the
browser for future sessions.

Wildcard also includes a formula language that enables more
sophisticated customizations. When traveling without a car, it's useful
to evaluate potential places to stay based on how walkable the
surroundings are. Using Wildcard formulas, the user can integrate Airbnb
with Walkscore, an API that rates the walkability of any location on a
1-100 scale. When they type in a \texttt{walkscore} formula with the
latitude and longitude of the listing, it sets the walk score as the
value of the cell. Because the cell's contents are injected into the
page, the score also shows up in the UI{
(Figure~\ref{fig:airbnb-demo}C)}.

\begin{figure}
\hypertarget{fig:expedia-demo}{%
\centering
\includegraphics{media/expedia-demo-300dpi.png}
\caption{Using Wildcard to augment the Expedia page for booking a
flight}\label{fig:expedia-demo}
}
\end{figure}

\hypertarget{customizing-a-form-input}{%
\subsection{Customizing a form input}\label{customizing-a-form-input}}

It might seem that Wildcard is only useful on websites that display
lists of tabular data like search results. But in fact, the table
metaphor is flexible enough to represent many types of data. For
example, a flight search form on Expedia can be represented as a single
row, with a column corresponding to each input. {
Figure~\ref{fig:expedia-demo} shows an overview of augmenting the
Expedia site.}

In previous examples the table cells were read-only, since users cannot
change the name or price of an Airbnb listing. In this case, the cells
are writable, which means that changes in the table are reflected in the
form inputs. This becomes especially useful when combined with GUI
widgets that can edit the value of a table cell.

Filling in dates for a flight search typically requires opening up a
separate calendar app to find the right dates and then manually copying
them into the form. In Wildcard, the user can avoid this cumbersome
workflow by using a datepicker widget that includes the user's personal
calendar events{ (Figure~\ref{fig:expedia-demo}B)}. The user can
directly click on the right date, and it gets inserted into the
spreadsheet and the original form.

Here we've presented just a few use cases for spreadsheet-driven
customization. We think the spreadsheet model is flexible enough to
support a wide range of other useful modifications, while remaining
familiar and easy to use.

\hypertarget{sec:implementation}{%
\section{System Implementation}\label{sec:implementation}}

Wildcard is written in Typescript. It is currently injected into pages
using the \href{https://www.tampermonkey.net/}{Tampermonkey} userscript
manager, but in the future we plan to deploy it as a standalone browser
extension to make it easier to install.

In order to promote extensibility, Wildcard is implemented as a small
core program along with several types of plugins: site adapters,
formulas, and cell editors. The core contains functionality for
displaying the data table and handling user interactions, and the table
implementation is built using the
\href{https://handsontable.com/}{Handsontable} Javascript library.

\begin{figure}
\hypertarget{fig:architecture}{%
\centering
\includegraphics{media/architecture-clean.png}
\caption{The architecture of the Wildcard
system}\label{fig:architecture}
}
\end{figure}

Site adapters are a key part of Wildcard, since they specify the
bidirectional connection between the web page and its structured data
representation.

Wildcard provides an interface for concisely expressing how the DOM
should be mapped to the spreadsheet table. For example, here is a code
snippet for extracting the name of an Airbnb listing:

\begin{Shaded}
\begin{Highlighting}[]
\OperatorTok{\{}
  \CommentTok{// Field metadata:}
\NormalTok{  fieldName}\OperatorTok{:} \StringTok{"name"}\OperatorTok{,} \CommentTok{// The name of the field}
\NormalTok{  type}\OperatorTok{:} \StringTok{"text"}\OperatorTok{,}      \CommentTok{// The type of the field}
\NormalTok{  readOnly}\OperatorTok{:} \KeywordTok{true}\OperatorTok{,}    \CommentTok{// Whether the user can edit the field}

  \CommentTok{// Function to extract DOM element:}
\NormalTok{  el}\OperatorTok{:}\NormalTok{ (row) }\KeywordTok{=>} \VariableTok{row}\OperatorTok{.}\FunctionTok{querySelector}\NormalTok{(}\VerbatimStringTok{\textasciigrave{}.}\SpecialCharTok{$\{}\NormalTok{titleClass}\SpecialCharTok{\}}\VerbatimStringTok{\textasciigrave{}}\NormalTok{)}\OperatorTok{,}
\OperatorTok{\}}
\end{Highlighting}
\end{Shaded}

Sometimes sophisticated scraping techniques are necessary to extract the
necessary data from the application. For example, when we have
prototyped mechanisms for site adapters to observe AJAX requests made by
the browser and extract data directly from JSON responses. This
mechanism was used to implement the Airbnb Walkscore example, since
latitude and longitude aren't shown in the Airbnb UI, but they are
available in AJAX responses. This technique seems promising because AJAX
responses tend to already contain data in a structured form, and web
applications increasingly load data using AJAX. Another technique we
might consider adding is the ability for site adapters to scrape data
across multiple pages for paginated lists of results (as explored in
\autocite{huynh2006}).

The site adapter also needs to support the reverse direction: sending
updates from the table to the original page. Most DOM manipulation is
not performed directly by the site adapter. Instead, Wildcard
automatically mutates the DOM to reflect the spreadsheet state, using
the same declarative spec shown above. The only exception is row actions
(like favoriting an Airbnb listing), which are implemented as imperative
Javascript functions that can can mutate the DOM, simulate clicks on
buttons, etc.

\hypertarget{sec:design-principles}{%
\section{Design principles}\label{sec:design-principles}}

The idea of spreadsheet-driven customization is guided by several design
principles, inspired by prior work and our own experimentation. We think
these principles can also broadly inform the design of tools for end
user software customization.

\hypertarget{expose-a-universal-data-structure}{%
\subsection{Expose a universal data
structure}\label{expose-a-universal-data-structure}}

Today, most personal computing consists of using applications, which
bundle together behavior and data to provide some set of functionality.
While there are some limited points of interoperability, applications
generally are designed to operate independently of one another.

Computing does not need to be organized this way. For example, UNIX
offers a compelling alternative design: many small single-purpose
utilities, all of which manipulate a universal format of text streams.
The universal format creates a high degree of leverage from tools: users
can get a lot of utility from deeply mastering a text editor and text
manipulation utilities like \texttt{cut} and \texttt{sed}, because these
tools can be applied to a huge variety of tasks. A user's preferred text
editor can even serve as an interactive input mechanism in shell
programs, e.g.~for editing git commit messages.

Spreadsheet-driven customization aims to bring some of the UNIX ethos to
the world of isolated Web applications, by creating a consistent data
structure to represent the data inside many applications. In UNIX, the
universal format is a text stream; in Wildcard, it is a relational
table. Because Wildcard maps the data from all applications to the table
format, users can invest in mastering the Wildcard table editor, the
formula language, and cell editor UIs, and reuse those same tools to
customize many different applications.

This idea relates to Beaudouin-Lafon and Mackay's notion of
\emph{polymorphic interface instruments} \autocite{beaudouin-lafon2000},
which are UI elements that can be used in different contexts (for
example, a color picker that can be used in many different drawing
applications). It also relates to ideas of literacy in a medium. diSessa
notes that textual literacy rests on the fact that writing can be
adapted to many different genres and uses \autocite{disessa2000}; if
people needed to relearn reading and writing from scratch when switching
from essays to emails, the medium would lose most of its potency. Having
generic tools is especially important for software customization,
because the most common barrier to customizing software is not having
enough time \autocite{mackay1991}. It's more likely that people will
customize software regularly if they can reuse the same tools across
many applications.

This design principle leads to several challenges. First, any universal
abstraction has its constraints, and can't necessarily naturally express
the data in every application. We plan to explore the limits of the
table abstraction further, by trying to build adapters for more sites
with varied data formats. We expect that many types of data can fit
easily into tables: lists of search results, news articles, and messages
can all naturally be seen as relations. For sites that use document
structures (e.g.~Google Docs) or graph structures (e.g.~social
networks), it may prove more challenging to map internal data to the
table abstraction.

\hypertarget{low-floor-high-ceiling}{%
\subsection{Low floor, high ceiling}\label{low-floor-high-ceiling}}

Seymour Papert advocated for programming systems to have a ``low
floor,'' making it easy for novices to get started, and a ``high
ceiling,'' providing a large range of possibilities for more
sophisticated users \autocite{resnick2016}. Our goal is for
spreadsheet-driven customization to meet both of these criteria.

One of the most interesting properties of spreadsheets is that users who
are only know a tiny sliver of their functionality (e.g., storing tables
of numbers or computing simple sums) can still use them in genuinely
valuable ways. The fact that useful tasks can be performed early on
supports the user's natural motivation to continue using the tool, and
to eventually learn its more powerful features if needed
\autocite{nardi1991}. In contrast, many traditional programming systems
require an enormous upfront time investment before someone is able to
write a program that helps them achieve a useful task in their life.

As part of ensuring a low floor, we have focused on including genuinely
valuable features for novices. For example, a user can sort a table with
a single click, or simply type in some annotations. We would expect many
Wildcard users to start by using these simpler features before
potentially moving on to more sophisticated features like formulas.

Another aspect of providing a low floor is providing an ``in-place
toolchain'' \autocite{inkandswitch2019}: minimizing the effort of moving
from using to customizing, by making customization tools available in
the same environment where the user is already using the software. This
quality is distinct from the level of technical skill needed to use the
tool. For example, setting up a workflow trigger in an end user
programming system like \href{https://ifttt.com/}{IFTTT} does not
require much technical skill, but does require leaving the user's normal
software and entering a separate environment; conversely, running a
Javascript snippet in the browser console requires programming skills,
but can be done immediately and casually in the flow of using a website.

\begin{longtable}[]{@{}llc@{}}
\toprule
\endhead
& \emph{In-place} & \emph{Not in-place}\tabularnewline
\emph{End user friendly} & \textbf{Wildcard} & IFTTT\tabularnewline
\emph{Requires programming} & browser JS console & forking open source
software\tabularnewline
\bottomrule
\end{longtable}

Wildcard provides an in-place toolchain because the spreadsheet can be
instantly opened in the browser window while using any supported
website. Once the user starts editing, Wildcard also provides live
feedback, so that even if a user isn't yet totally familiar with
Wildcard, they can learn to use the system through experimentation.

Since we have still only built several site adapters and demos, it is
still too early to tell exactly how high the ceiling is for the
customizations that can be achieved with Wildcard. But we think that
with enough operators, the formula language could support a wide variety
of customizations---people have managed to use simple spreadsheet
formula languages to solve a surprisingly large range of problems. We
plan to explore this aspect further by trying to solve more real
problems with the system and observing where limitations emerge in
practice.

\hypertarget{build-for-multiple-tiers-of-users}{%
\subsection{Build for multiple tiers of
users}\label{build-for-multiple-tiers-of-users}}

Real-world spreadsheet usage is highly collaborative. Many users just
perform simple changes, while their coworkers help with writing more
complex formulas or even programming macros \autocite{nardi1990}.
Inspired by this, we aim to make spreadsheet-driven customization a
collaborative activity that combines the different abilities of
different users.

The main way we do this is by separating website customization into two
separate stages: structured data extraction, performed by programmers
who code site adapters in Javascript, and then customization using the
resulting spreadsheet, which is available to all end users. This
architecture frees end users from needing to think about data
extraction, and enables a community of end users to reuse the efforts of
programmers building site adapters.

The group of users building adapters does not necessarily need to be
limited only to programmers. In the future, we plan to explore enabling
end users to also create site adapters, drawing on related work in this
area \autocite{chasins2018,huynh2006}. But even then, we still envision
a separation between highly motivated, tech-savvy end users building
adapters, and more casual end users just using the spreadsheet view.

Another group to consider is the first party developers of the original
software. Spreadsheet-driven customization does not depend on
cooperation from first-party website developers, but if they were to
expose structured data in their web application clients, it would
eliminate the need for third party site adapters. We think there are
compelling reasons for first parties to consider doing this. Providing
Wildcard support would allow users to build extensions to fulfill their
own feature requests. It also would not necessarily require much effort,
since adding Wildcard support could be fairly straightforward for a
first-party that already has direct access to the structured data in the
page. There is also precedent for first parties implementing an official
client extension API: for several years, Google maintained an official
extension API in Gmail for Greasemonkey scripts to use.\footnote{Incidentally,
  since then, third parties have continued to maintain stable Gmail
  extension APIs used by many browser extensions
  \autocite{streak,talwar2019}, illustrating the potential of
  collaboratively maintaining third party adapters.}

\hypertarget{sec:related-work}{%
\section{Related work}\label{sec:related-work}}

\hypertarget{malleable-software}{%
\subsection{Malleable software}\label{malleable-software}}

In the broadest sense, Wildcard is inspired by systems aiming to make
software into a dynamic medium where end users frequently create and
modify software to meet their own needs, rather than only consuming
applications built by programmers. These systems include Smalltalk
\autocite{kay1977}, Hypercard \autocite{hypercard2019} , Boxer
\autocite{disessa1986}, Webstrates \autocite{klokmose2015}, and
Dynamicland \autocite{victor}.\footnote{The project's name Wildcard
  comes from the internal pre-release name for Hypercard, which doubly
  inspired our work by promoting both software modification by end users
  and the ideas behind the Web.}

While similar in broad goals, Wildcard employs a different solution
strategy. These projects generally require building software from
scratch in a new environment, whereas Wildcard aims to maximize the
malleability of software built with existing tools.

With substantial future work, we think Wildcard could become more
similar to these other projects, growing from a platform for tweaking
existing software into a platform for building new software from
scratch. This would likely end up resembling existing tools for building
spreadsheet-driven applications (discussed more below), but with an
additional focus on customization by end users of the software.

\hypertarget{web-customization}{%
\subsection{Web customization}\label{web-customization}}

Wildcard's goals are closely shared with other systems that provide
interfaces in the browser for end users to customize websites.

\hypertarget{structured-augmentation}{%
\subsubsection{Structured augmentation}\label{structured-augmentation}}

Wildcard's solution approach is most similar to other tools that
identify structured data in a web page, and then use that structure to
support end user customization of the page.

Sifter \autocite{huynh2006} enables users to sort and filter lists of
data on web pages, which resembles Wildcard's sorting functionality. The
underlying mechanism is similar, since Sifter extracts structured data
from the page to enable its user-facing functionality. However, the
systems also have significant differences:

\begin{itemize}
\tightlist
\item
  Wildcard aims support a broad set of customizations, but Sifter only
  supports sorting and filtering.
\item
  Wildcard shows the structured data table directly to the user. Sifter
  only shows sorting controls, and does not reveal the underlying data
  table.
\item
  Wildcard hides data extraction from end users. Sifter uses a
  combination of automated heuristics and interactive user feedback to
  involve end users in extracting data.
\end{itemize}

Thresher \autocite{hogue2005} enables users to create wrappers which map
unstructured website content to Semantic Web content. Like Wildcard and
Sifter, Thresher augments the page; once semantic content has been
identified, it creates context menus in the original website which allow
users to take actions based on that content. Wildcard and Thresher focus
on complementary parts of the customization process. Thresher aims to
enable end users to create content wrappers, but the actions available
on the structured data are determined in advance. Conversely, Wildcard
delegates wrapper creation to programmers, but gives end users more
flexibility to use the structured data in an open-ended way.

\hypertarget{sloppy-augmentation}{%
\subsubsection{Sloppy augmentation}\label{sloppy-augmentation}}

``Sloppy programming'' \autocite{little2010} tools like Chickenfoot
\autocite{bolin2005} and Coscripter \autocite{leshed2008} enable users
to create web automation scripts without directly interacting with the
DOM. The scripts can perform actions like filling in text boxes and
clicking on buttons. Users express the desired page elements in natural,
informal terms (e.g.~writing ``the username box'' to represent the
textbox closest to the label ``username''), and then the system uses a
set of heuristics to determine which elements most likely match the
user's intent. This approach allows for expressing a wide variety of
commands with minimal training, but it also has downsides
\autocite{little2010}:

\begin{itemize}
\tightlist
\item
  \emph{Reliability}: It can be difficult to know whether a command will
  consistently work over time. In addition to changes to the website,
  changes to the heuristics can also cause problems.
\item
  \emph{Discoverability}: it can be difficult for users to discover the
  space of possible commands.
\end{itemize}

Wildcard offers a sharp contrast to sloppy programming, instead choosing
to expose a high degree of structure through the familiar spreadsheet
table. Wildcard offers more consistency; for example, clicking a sort
header will always work correctly as long as the site adapter is
maintained. Wildcard also offers clearer affordances for what types of
actions are possible, or, crucially, what actions are \emph{not}
possible, which is useful to know. On the other hand, Wildcard's
explicit site adapter approach means that fewer websites can be
customized, and also that users can't perform customizations if the
relevant data is not in the spreadsheet.

\hypertarget{spreadsheet-based-app-builders}{%
\subsection{Spreadsheet-based app
builders}\label{spreadsheet-based-app-builders}}

Many others have had the powerful realization that a spreadsheet can
serve as an backing data store and computation layer for an interactive
web application, enabling end users to create such applications more
easily. Research projects like Object Spreadsheets
\autocite{mccutchen2016}, Quilt \autocite{benson2014}, Gneiss
\autocite{chang2014}, and Marmite \autocite{wong2007}, as well as
commercial tools like Airtable \autocite{zotero-79} and Glide
\autocite{zotero-81} allow users to view data in a spreadsheet table,
compute over the data using formulas, and then connect the table to a
GUI. Because many users are already familiar with using spreadsheets,
this way of creating applications tends to be much easier than
traditional software methods; for example, in a user study of Quilt,
many users were able to create applications in under 10 minutes, even if
they expected it would take them many hours.

Wildcard builds on this idea, but applies it to modifying existing
applications, rather than building new applications from scratch. For
many people, we suspect that tweaking existing applications provides
more motivation as a starting point for programming than creating a new
application from scratch.

An important design decision for tools in this space is how to deviate
from traditional spreadsheets like Microsoft Excel or Google Sheets.
Quilt and Glide use existing spreadsheet software as a backend,
providing maximum familiarity for users, and even compatibility with
existing spreadsheets. Gneiss has its own spreadsheet implementation
with additional features useful for building GUIs. Marmite provides a
live data view that resembles a spreadsheet, but programming is actually
done using a separate data flow pane rather than spreadsheet formulas.
(Marmite's approach led to some confusion in a user study, because users
expected behavior more similar to spreadsheets \autocite{wong2007}.)
Airtable deviates the furthest: although the user interface resembles a
spreadsheet, the underlying structure is a relational database with
typed columns. Wildcard's table is most similar to Airtable; the
structure of a relational table is most appropriate for most data in
websites, and we have not yet found a need for arbitrary untyped cells.

\hypertarget{web-scraping-data-extraction}{%
\subsection{Web scraping / data
extraction}\label{web-scraping-data-extraction}}

Web scraping tools focus on extracting structured data out of
unstructured web pages. Web scraping is closely related to the
implementation of Wildcard, but has different end goals: web scraping
generally extracts static data for processing in another environment,
whereas Wildcard customizes the application's UI by maintaining a
bidirectional connection between the extracted data and the page.

Web scraping tools differ in how much structure they attempt to map onto
the data. Some tools like Rousillon \autocite{chasins2018} extract data
in a minimally structured relational format; other tools like Piggy Bank
\autocite{huynh2005} more ambitiously map the data to a rich semantic
schema. In Wildcard, we chose to avoid semantic schemas, in order to
minimize the work associated with creating a site adapter.

In the future, we might try to integrate existing web scraping tools, to
help create more reliable site adapters for Wildcard with less work, and
to open up adapter creation to end users.

\hypertarget{future-work}{%
\section{Future work}\label{future-work}}

There are still many open questions about spreadsheet-driven
customization which we hope to answer through targeted development and
usage of the Wildcard prototype.

The most important question is: what are the limits of this
computational model? What types of useful customizations can it support
or not support? While initial demos suggest a variety of use cases, we
plan to develop more site adapters and demos to explore this question
further. We will start with privately testing the system with our own
needs, and then eventually deploy the tool publicly, once the API is
stable enough and can support a critical mass of sites and use cases. We
also plan to run usability studies to evaluate and improve the design of
the tool.

We suspect that two areas of the current model may prove limiting.
First, Wildcard's data model currently shows a single table at a time,
without any notion of relationships between tables. A richer data model
with foreign keys might help support certain use cases. For designing a
spreadsheet interface on top of a richer relational model, we could
learn from other systems with this design
\autocite{mccutchen2016,bakke2016}.

A second limitation is that there is currently no way for end users to
express imperative workflows with sequences of actions. Related work has
shown these types of workflows to be useful
\autocite{bolin2005,leshed2008} in practice. It's not clear how such
workflows could fit into Wildcard, since spreadsheets have a
fundamentally different computation model, and site adapters created by
programmers cannot easily account for every possible action a user would
want to take in a page.

Another open question is how easily site adapters can be created and
maintained for real websites, which often include complex markup and
change frequently. We plan to explore this question by creating adapters
for sites with different data structures and in different domains, and
perhaps running user tests with programmers. Possible future
improvements we've considered include developing automated heuristics to
assist with the adapter creation process, and developing new
abstractions that make it easier for programmers to efficiently create
new robust adapters.

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

In this paper, we have presented \emph{spreadsheet-driven
customization}, a technique that enables end users to customize software
by augmenting an application's UI with a spreadsheet. We hope that this
technique contributes to making the Web into a more dynamic medium that
users can mold to their own needs.

We plan to continue developing the Wildcard prototype and to eventually
deploy it as an open-source tool. To receive future updates on Wildcard
and notifications about a public release, \href{}{sign up for the email
newsletter}.

We are also looking for private beta testers. If you have an idea for
how you might want to use Wildcard, please
\href{mailto:glitt@mit.edu}{get in touch}. We would love to hear about
your needs and help find ways to use Wildcard to solve them.

\acks
Thanks to Glen Chiacchieri, David Karger, Steve Krouse, Rob Miller, Arvind Satyanarayan, Daniel Windham, and Maggie Yellen for providing valuable feedback on this work.

\printbibliography

\end{document}

% Local Variables:
% TeX-engine: luatex
% End: