### 2020-01-14 paper.tex
done

% -*- coding: utf-8; -*-
% vim: set fileencoding=utf-8 :
\documentclass[english,submission]{programming}
%% First parameter: the language is 'english'.
%% Second parameter: use 'submission' for initial submission, remove it for camera-ready (see 5.1)

\usepackage{longtable,booktabs}
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\newenvironment{Shaded}{}{}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.49,0.56,0.16}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.53,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.73,0.13,0.13}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.40,0.40,0.40}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.74,0.48,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.73,0.40,0.53}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.10,0.09,0.49}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}

\usepackage[backend=biber]{biblatex}
\addbibresource{references2.bib}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\begin{document}

\title{Wildcard: Spreadsheet-Driven Customization of Web Applications}
\subtitle{}% optional

\author{Geoffrey Litt}
\authorinfo{is a PhD student at MIT in the Software Design Group. His research focuses on programming languages and environments, particularly tools for end user programming and live programming. Contact him at \email{glitt@mit.edu}.}
\author{Daniel Jackson}
\authorinfo{is Professor in the Department of Electrical Engineering and Computer Science at MIT,  a MacVicar teaching fellow, and Associate Director of the Computer Science and Artificial Intelligence Laboratory. He leads the Software Design Group. Contact him at \email{dnj@csail.mit.edu}.}
\affiliation{Massachusetts Institute of Technology}

\keywords{end-user programming, software customization, web browser extensions} % please provide 1--5 keywords

%%%%%%%%%%%%%%%%%%
%% These data MUST be filled for your submission. (see 5.3)
\paperdetails{
  %% perspective options are: art, sciencetheoretical, scienceempirical, engineering.
  %% Choose exactly the one that best describes this work. (see 2.1)
  perspective=art,
  %% State one or more areas, separated by a comma. (see 2.2)
  %% Please see list of areas in http://programming-journal.org/cfp/
  %% The list is open-ended, so use other areas if yours is/are not listed.
  area={Programming environments, Visual and live programming},
  %% You may choose the license for your paper (see 3.)
  %% License options include: cc-by (default), cc-by-nc
  % license=cc-by-sa,
}
%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%
%% These data are provided by the editors. May be left out on submission.
%\paperdetails{
%  submitted=2016-08-10,
%  published=2016-10-11,
%  year=2016,
%  volume=1,
%  issue=1,
%  articlenumber=1,
%}
%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Please go to https://dl.acm.org/ccs/ccs.cfm and generate your Classification
% System [view CCS TeX Code] stanz and copy _all of it_ to this place.
%% From HERE
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10011007.10011006.10011066.10011069</concept_id>
<concept_desc>Software and its engineering~Integrated and visual development environments</concept_desc>
<concept_significance>500</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Software and its engineering~Integrated and visual development environments}
% To HERE
%%%%%%%%%%%%%%%%%%%%%%%

\maketitle

% Please always include the abstract.
% The abstract MUST be written according to the directives stated in 
% http://programming-journal.org/submission/
% Failure to adhere to the abstract directives may result in the paper
% being returned to the authors.
\begin{abstract}
Many Web applications do not meet the precise needs of their users.
Browser extensions offer a way to customize web applications, but most
people do not have the programming skills to implement their own
extensions.

In this paper, we present \emph{spreadsheet-driven customization}, a
technique that enables end users to customize software without doing any
traditional programming. The idea is to augment an application's UI with
a spreadsheet that is synchronized with the application's data. When the
user manipulates the spreadsheet, the underlying data is modified and
the changes are propagated to the UI, and vice versa.

We have implemented this technique in a prototype browser extension
called Wildcard. Through concrete examples, we demonstrate that Wildcard
can support useful customizations---ranging from sorting lists of search
results to showing related data from web APIs---on top of existing
websites. We also present the design principles underlying our
prototype.

Customization can lead to dramatically better experiences with software.
We think that spreadsheet-driven customization offers a promising new
approach to unlocking this benefit for all users, not just programmers.

This paper is best viewed in the online version (\url{https://www.geoffreylitt.com/wildcard/salon2020}), which includes videos demonstrating the interface.
\end{abstract}

\hypertarget{sec:introduction}{%
\section{Introduction}\label{sec:introduction}}

Web applications often don't meet the precise needs of their users.
Sometimes there is a browser extension available to patch an issue, and
if the user is a programmer they might be able to fix it themselves. But
for most people, the only recourse is to complain to the developers, or,
more likely, to simply give up. Back in 1977, in \emph{Personal Dynamic
Media} \autocite{kay1977}, Alan Kay envisioned personal computing as a
medium that let a user ``mold and channel its power to his own needs,''
but today, software behaves more like concrete than clay.

In this paper, we present \emph{spreadsheet-driven customization}, a
technique that enables end users to customize software without doing any
traditional programming. The idea is to augment an application's UI with
a spreadsheet that is synchronized with the application's data. When the
user manipulates the spreadsheet, the underlying data is modified and
the changes are propagated to the UI, and vice versa.

We have implemented this technique in a prototype browser extension
called Wildcard and used it to build demos which suggest that this
paradigm can support useful customizations, ranging from sorting lists
of data to adding whole new features to applications (shown in
Section~\ref{sec:demos}).

\begin{figure}
\hypertarget{fig:overview}{%
\centering
\includegraphics{media/overview-300dpi.png}
\caption{An overview of spreadsheet-driven
customization}\label{fig:overview}
}
\end{figure}

Our approach requires extracting structured data from the user
interfaces of existing applications, but we hide the complexity of data
extraction from end users. Programmers write \emph{site adapters} which
use web scraping techniques to extract structured data from existing
applications and map it to the spreadsheet table. Our prototype suggests
that it is possible to implement site adapters for real websites;
Section~\ref{sec:implementation} describes some of the techniques and
challenges involved.

Spreadsheet-driven customization is based on three design principles,
described in Section~\ref{sec:design-principles}:

\begin{itemize}
\tightlist
\item
  \textbf{Expose a universal structure}: By mapping the data in all
  applications to the same universal table abstraction, Wildcard enables
  users to learn a generic customization toolkit that they can apply to
  any site.
\item
  \textbf{Low floor, high ceiling}: Wildcard provides an easy entry
  point for end users, since small tweaks like sorting data can be
  performed with a single click. At the same time, it also supports a
  variety of richer customizations, like adding private annotations to a
  webpage or joining in related data from a web API.
\item
  \textbf{Design for an ecosystem of users}: Spreadsheet-driven
  customization combines the efforts of programmers and end users,
  rather than putting the full burden on end users. We also envision a
  role for first party developers to help make their applications more
  customizable.
\end{itemize}

Prior work \autocite{mccutchen2016,benson2014,chang2014} has enabled end
users to create ``spreadsheet-driven applications'' which use
spreadsheets as a backing data layer. Spreadsheet-driven
\emph{customization} applies this idea in a different context:
customizing existing software, rather than building new software from
scratch. Our technique does not require that the application actually be
backed by a spreadsheet; it merely uses the spreadsheet as an interface
for viewing and modifying the internal state of the application. In
Section~\ref{sec:related-work}, we describe further how Wildcard relates
to existing work on spreadsheet-driven apps, as well as work in other
areas like web customization and web scraping.

The Wildcard extension is currently an early research prototype. We plan
to continue testing the system with our own use cases to explore how the
spreadsheet abstraction maps to real websites and customization needs.
Eventually we plan to release the tool publicly in order to learn from
real use cases, discover usability challenges, and to test the
feasibility of building and maintaining site adapters.
Section~\ref{sec:conclusion} describes open questions we plan to
continue exploring, as well as our future plans.

\hypertarget{sec:demos}{%
\section{Demos}\label{sec:demos}}

Here are some examples of using Wildcard to customize websites in useful
ways.{ These demos are best viewed as videos in the online version of
this paper (\url{https://www.geoffreylitt.com/wildcard/salon2020}).}

\begin{figure}
\hypertarget{fig:airbnb-demo}{%
\centering
\includegraphics{media/airbnb-demo-300dpi.png}
\caption{Using Wildcard to augment the Airbnb search page for booking
accommodations}\label{fig:airbnb-demo}
}
\end{figure}

\hypertarget{augmenting-search-results}{%
\subsection{Augmenting search results}\label{augmenting-search-results}}

In 2012, the travel site Airbnb removed the ability to sort
accommodation searches by price. Users could still filter by price
range, but could no longer view the cheapest listings first. Many users
complained that the change seemed hostile to users. ``It's so
frustrating! What is the logic behind not having this function?'' said
one user on the
\href{https://community.withairbnb.com/t5/Hosting/Sorting-listing-by-price/td-p/559404}{Airbnb
support forum}. Alas, the feature remains missing to this day.

Using Wildcard, the user can fix this omission, while leaving the page's
design and the rest of its functionality unchanged.{
Figure~\ref{fig:airbnb-demo} shows an overview of augmenting the Airbnb
site.} First, the user opens up the Wildcard panel, which shows a table
corresponding to the search results in the page. As they click around in
the table, the corresponding row in the page is highlighted to indicate
the mapping between the views.

Then, the user clicks on the price column header to sort the spreadsheet
and the Airbnb UI by price{ (Figure~\ref{fig:airbnb-demo}, Note A)}.
They also filter to listings with a user rating above 4.5 (another
feature missing in the original Airbnb UI).

After manipulating the data, the user closes the table view and
continues using the website. Because the application's UI usually has a
nicer visual design than a spreadsheet, Wildcard does not aim to replace
it---at any time, the user can use either the UI, the spreadsheet, or
both together.

Many websites that show lists of data also offer actions on rows in the
table, like adding an item to a shopping cart. Wildcard has the ability
to make these ``row actions'' available in the data table through the
site adapter. In the Airbnb UI, saving multiple listings to a Favorites
list requires tediously clicking through them one by one. Using Wildcard
row actions, the user can select multiple rows and favorite all of them
with a single click{ (Figure~\ref{fig:airbnb-demo}, Note B)}. Similarly,
the user can open the detailed pages for many listings at once.

Next, the user wants to jot down some notes about each listing. To do
this, they type some notes into an additional column in each row, and
the notes appear inside the listings in the original UI{
(Figure~\ref{fig:airbnb-demo}, Note C)}. The annotations are saved in
the browser and associated with the backend ID of the listing, so they
will appear in future browser sessions that display the same listing.

Wildcard also includes a formula language that enables more
sophisticated customizations. When traveling without a car, it's useful
to evaluate potential places to stay based on how walkable the
surroundings are. Using a formula, the user can integrate Airbnb with
Walkscore, an API that rates the walkability of any location on a 1-100
scale. When the user calls the \texttt{walkscore} formula with the
listing's latitude and longitude as arguments, it returns the walk score
for that location and shows it as the cell value. Because the cell's
contents are injected into the page, the score also appears in the UI{
(Figure~\ref{fig:airbnb-demo}, Note D)}.

\hypertarget{snoozing-todos}{%
\subsection{Snoozing todos}\label{snoozing-todos}}

\begin{figure}
\hypertarget{fig:todomvc-demo}{%
\centering
\includegraphics{media/todomvc-demo-300dpi.png}
\caption{Using Wildcard to add a ``snooze'' feature to the TodoMVC todo
list app}\label{fig:todomvc-demo}
}
\end{figure}

In addition to fetching data from other sources, Wildcard formulas can
also perform computations. In this example, the user would like to
augment the TodoMVC todo list app with a ``snooze'' feature, which will
temporarily hide a todo from the list until a certain date.{
Figure~\ref{fig:todomvc-demo} shows an overview of this customization.}

The user opens the table view, which shows the text and completed status
of each todo. They start the customization by adding a new column to
store the snooze date for each todo{ (Figure~\ref{fig:todomvc-demo},
Note A)}.

The next step is to hide snoozed todos. The user creates a
\texttt{snoozed?} column, which uses a formula to compute whether a todo
is snoozed---i.e., whether it has a snooze date in the future{
(Figure~\ref{fig:todomvc-demo}, Note B)}. Then, they simply filter the
table to hide the snoozed todos{ (Figure~\ref{fig:todomvc-demo}, Note
C)}.

Because the built-in \texttt{NOW()} function always returns the current
datetime, snoozed todos will automatically appear once their snooze date
arrives.

Because this implementation of snoozing was built on the spreadsheet
abstraction, it is completely decoupled from this particular todo list
app. We envision that users could share these types of customizations as
generic browser extensions, which could be applied to any site supported
by Wildcard with no additional effort.

\hypertarget{adding-a-custom-datepicker}{%
\subsection{Adding a custom
datepicker}\label{adding-a-custom-datepicker}}

\begin{figure}
\hypertarget{fig:expedia-demo}{%
\centering
\includegraphics{media/expedia-demo-300dpi.png}
\caption{Using Wildcard to augment the Expedia page for booking a
flight}\label{fig:expedia-demo}
}
\end{figure}

It might seem that Wildcard is only useful on websites that display
lists of tabular data, but the table metaphor is flexible enough to
represent many types of data. For example, a flight search form on
Expedia can be represented as a single row, with a column corresponding
to each input{ (Figure~\ref{fig:expedia-demo}, Note A)}.

In some of the previous examples, the table cells were read-only
(because users can't, for example, change the name or price of an Airbnb
listing). In this case, the cells are writable, which means that changes
in the table are reflected in the form inputs. This becomes especially
useful when combined with GUI widgets that can edit the value of a table
cell.

Filling in dates for a flight search often requires a cumbersome
workflow: open up a separate calendar app, find the dates for the trip,
and then carefully copy them into the form. In Wildcard, the user can
avoid this by using a datepicker widget that shows the user's personal
calendar events{ (Figure~\ref{fig:expedia-demo}, Note B)}. The user can
directly click on the correct date, and it gets inserted into both the
spreadsheet and the original form.

In this section we've presented just a few use cases for
spreadsheet-driven customization, to suggest some possibilities of the
paradigm. We think the spreadsheet model is flexible enough to support a
wide range of other useful modifications while remaining familiar and
easy to use.

\hypertarget{sec:implementation}{%
\section{System Implementation}\label{sec:implementation}}

Wildcard is written in Typescript. It is currently injected into pages
using the \href{https://www.tampermonkey.net/}{Tampermonkey} userscript
manager, but in the future we plan to deploy it as a standalone browser
extension to make it easier to install.

In order to promote extensibility, Wildcard is implemented as a small
core program along with several types of plugins: site adapters,
formulas, and cell editors. The core contains functionality for
displaying the data table and handling user interactions, and is built
on the \href{https://handsontable.com/}{Handsontable} Javascript
library.

\begin{figure}
\hypertarget{fig:architecture}{%
\centering
\includegraphics{media/architecture-300dpi.png}
\caption{The architecture of the Wildcard
prototype}\label{fig:architecture}
}
\end{figure}

Site adapters are a key part of Wildcard, since they specify the
bidirectional connection between the web page and its structured data
representation. Wildcard represents data in a relational table with
named and typed columns, and provides an interface for concisely
expressing how the DOM should be mapped to this structure. For example,
here is a code snippet for extracting the name of an Airbnb listing.

\begin{Shaded}
\begin{Highlighting}[]
\OperatorTok{\{}
  \CommentTok{// Field metadata:}
\NormalTok{  fieldName}\OperatorTok{:} \StringTok{"name"}\OperatorTok{,}
\NormalTok{  type}\OperatorTok{:} \StringTok{"text"}\OperatorTok{,}
\NormalTok{  readOnly}\OperatorTok{:} \KeywordTok{true}\OperatorTok{,}

  \CommentTok{// Function to extract DOM element:}
\NormalTok{  el}\OperatorTok{:}\NormalTok{ (row) }\KeywordTok{=>} \VariableTok{row}\OperatorTok{.}\FunctionTok{querySelector}\NormalTok{(}\StringTok{".\_1ebt2xej"}\NormalTok{)}\OperatorTok{,}
\OperatorTok{\}}
\end{Highlighting}
\end{Shaded}

Sometimes, more advanced scraping techniques are necessary to extract
data from the application. For example, we have prototyped a mechanism
for site adapters to observe AJAX requests made by the browser and
extract data directly from JSON responses. This mechanism was used to
implement the Airbnb Walkscore example, since latitude and longitude
aren't shown in the Airbnb UI, but they are available in AJAX responses.
This technique seems promising because AJAX responses already contain
data in a structured form, and web applications are increasingly loading
data using AJAX. Another technique we might consider is enabling site
adapters to scrape data across multiple pages for paginated lists of
results (as explored in \autocite{huynh2006}).

Because Wildcard requires a live bidirectional connection between the UI
and the structured data, there are some additional challenges beyond
those typically associated with web scraping.

First, the site adapter needs to support sending updates from the table
to the original page. We have made this fairly straightforward, because
most DOM manipulation is not performed directly by the site adapter.
Instead, Wildcard automatically mutates the DOM to reflect the
spreadsheet state, using the same specification that was used to extract
the data. The only exception to this is row actions (like favoriting an
Airbnb listing), which are implemented as Javascript functions that can
perform arbitrary behaviors like simulating clicks on buttons or
mutating the DOM.

Another challenge is triggering updates to the spreadsheet data in
response to UI changes that happen after initial page load. Site
adapters are responsible for recognizing these changes by observing the
DOM. So far, we have been able to use event listeners and the
MutationObserver API to successfully observe changes, but it may prove
challenging to observe changes on some sites only through the DOM.

\hypertarget{sec:design-principles}{%
\section{Design Principles}\label{sec:design-principles}}

The idea of spreadsheet-driven customization is guided by three design
principles, inspired by prior work and our own experimentation. We think
these principles might also broadly inform the design of tools for end
user software customization.

\hypertarget{expose-a-universal-data-structure}{%
\subsection{Expose a universal data
structure}\label{expose-a-universal-data-structure}}

Today, most personal computing consists of using applications, which
bundle together behavior and data to provide some set of functionality.
While there are some limited points of interoperability, applications
generally are designed to operate independently of one another.

Computing does not need to be organized this way. For example, UNIX
offers a compelling alternative: many small single-purpose programs, all
of which manipulate a universal format of text streams. Users get high
leverage from mastering a text editor and other text manipulation tools,
since they can be applied to a huge variety of tasks. A user's preferred
text editor can even serve as an interactive input mechanism in shell
programs, e.g.~for editing git commit messages.

Spreadsheet-driven customization aims to bring some of this UNIX ethos
to the world of isolated Web applications by creating a consistent data
structure to represent the data inside many applications. In UNIX, the
universal format is a text stream; in Wildcard, it is a relational
table. Because Wildcard maps the data from all applications to the table
format, users can master the tools provided by Wildcard and apply them
to customize many different applications.

This idea relates to Beaudouin-Lafon and Mackay's notion of
\emph{polymorphic interface instruments} \autocite{beaudouin-lafon2000},
which are UI elements that can be used in different contexts (for
example, a color picker that can be used in many drawing applications).
It also relates to ideas of literacy in a medium. DiSessa notes that
textual literacy depends on the fact that writing can be adapted to many
different genres \autocite{disessa2000}: if people needed to relearn
reading and writing from scratch when switching from essays to emails,
the medium would lose its potency. Generic tools are especially
important for software customization, because a common barrier to
customizing software is not having enough time
\autocite{mackay1991}---it's more likely that people will customize
software frequently if they can reuse the same tools across many
applications.

This design principle leads to several challenges. First, any universal
abstraction has its limits, and may not naturally express the data in
every application. We plan to explore the limits of the table
abstraction further by trying to build adapters for more sites with
varied data formats. We expect that many types of data will fit easily
into tables: lists of search results, news articles, and messages can
all naturally be seen as relations. Documents (e.g.~Google Docs) or
graphs (e.g.~social network friend graphs) may prove more challenging to
map to the table abstraction.

\hypertarget{low-floor-high-ceiling}{%
\subsection{Low floor, high ceiling}\label{low-floor-high-ceiling}}

Seymour Papert advocated for programming systems to have a ``low
floor,'' making it easy for novices to get started, and a ``high
ceiling,'' providing a large range of possibilities for more
sophisticated users \autocite{resnick2016}. Our goal is for
spreadsheet-driven customization to meet both of these criteria.

One of the most interesting properties of spreadsheets is that users
familiar with only a tiny sliver of their functionality (e.g., storing
tables of numbers and computing simple sums) can still use them in
valuable ways. This supports the user's natural motivation to continue
using the tool, and to eventually learn its more powerful features if
needed \autocite{nardi1991}. In contrast, many traditional programming
systems require an enormous upfront time investment before someone is
able to write a program that helps them achieve a useful task.

As part of ensuring a low floor, we have focused on including features
that have immediate value for novices. For example, a user can sort a
table with a single click, or simply type in some annotations. We would
expect many Wildcard users to start by using these simpler features
before moving on to more sophisticated features like formulas.

Another aspect of providing a low floor is providing an ``in-place
toolchain'' \autocite{2019}: minimizing the effort of moving from using
to customizing by making customization tools available in the same
environment where the user is already using the software. This quality
is distinct from the level of technical skill needed to use the tool.
For example, setting up a workflow in an end user programming system
like \href{https://ifttt.com/}{IFTTT} does not require much programming
skill, but it does require leaving the user's normal software and
entering a separate environment; conversely, running a Javascript
snippet in the browser console requires programming skills, but can be
done immediately in the flow of using a website.

\begin{longtable}[]{@{}llc@{}}
\toprule
\endhead
& \emph{In-place} & \emph{Not in-place}\tabularnewline
\emph{End user friendly} & \textbf{Wildcard} & IFTTT\tabularnewline
\emph{Requires programming} & browser JS console & using Web APIs in
Python\tabularnewline
\bottomrule
\end{longtable}

Wildcard provides an in-place toolchain because the spreadsheet can be
instantly opened in the browser window while using any supported
website. Once the user starts editing, Wildcard also provides live
feedback, so that even if a user isn't yet totally familiar with
Wildcard, they can learn to use the system through experimentation.

Since we have only built a few site adapters and demos so far, it is
still too early to tell how high the ceiling is for the customizations
that can be achieved with Wildcard. But we think that with enough
built-in formulas (both standard spreadsheet formulas for computation,
and formulas for fetching data from other sources), the language could
support a wide variety of customizations. We plan to explore this aspect
further by trying to solve real problems and observing where limitations
emerge in practice (discussed more in Section~\ref{sec:conclusion}).

\hypertarget{design-for-an-ecosystem-of-users}{%
\subsection{Design for an ecosystem of
users}\label{design-for-an-ecosystem-of-users}}

Real-world spreadsheet usage is highly collaborative. Many users only
perform simple changes, while their coworkers help with writing complex
formulas or programming macros \autocite{nardi1990}. Inspired by this
collaborative dynamic, we aim for spreadsheet-driven customization to
bring together the abilities of different users.

The main way we do this is by separating website customization into two
separate stages: structured data extraction, performed by programmers
who code site adapters in Javascript, and then customization using the
resulting spreadsheet, which is available to all end users. This
architecture frees end users from needing to think about data
extraction, and enables a community of end users to reuse the efforts of
programmers building site adapters.

The group of users building adapters does not necessarily need to be
limited only to programmers. In the future, we plan to explore enabling
end users to also create site adapters, drawing on related work in this
area \autocite{chasins2018,huynh2006}. But even then, we still envision
a separation between highly motivated, tech-savvy end users building
adapters, and more casual end users just using the spreadsheet view.

The first party developers of the original software can also play a role
in enabling customization. Although spreadsheet-driven customization
does not depend on cooperation from first-party website developers,
their participation in exposing structured data would eliminate the need
for third party site adapters. The first parties could themselves
benefit from such a system, since providing Wildcard support---a fairly
low-effort undertaking for a developer with direct access to the
structured data---would allow users to build extensions to fulfill some
of their own feature requests. There is also precedent for first parties
implementing an official client extension API: for several years, Google
maintained an official extension API in Gmail for Greasemonkey scripts
to use.

Google no longer maintains its own API, but third parties have continued
to maintain successful replacements---Gmail.js \autocite{talwar2019} is
an open source wrapper API with over 70 contributors and hundreds of
commits. This library demonstrates another benefit of designing for an
ecosystem: it's easier for many programmers to collectively maintain a
generic site adapter that can support many extensions, rather than each
programmer creating custom scraping logic for their own single-purpose
extension.

\hypertarget{sec:related-work}{%
\section{Related Work}\label{sec:related-work}}

Spreadsheet-driven customization relates to existing work in four areas:
malleable software, web customization, spreadsheet-driven applications,
and web scraping.

\hypertarget{malleable-software}{%
\subsection{Malleable software}\label{malleable-software}}

In the broadest sense, Wildcard is inspired by systems aiming to make
software into a dynamic medium where end users frequently create and
modify software to meet their own needs, rather than only consuming
applications built by programmers. These systems include Smalltalk
\autocite{kay1977}, Hypercard \autocite{hypercard2019}, Boxer
\autocite{disessa1986}, Webstrates \autocite{klokmose2015}, and
Dynamicland \autocite{victor}. In fact, the name Wildcard comes from the
internal pre-release name for Hypercard, a project which doubly inspired
our work by promoting both software modification by end users and the
ideas behind the Web.

Although Wildcard shares broad goals with these projects, it employs a
different solution strategy. These other systems generally require
building software from scratch in a new environment. On the other hand,
Wildcard aims to maximize the malleability of software built with
existing tools.

With substantial future work, we think Wildcard could become more
similar to these other projects, growing from a platform for tweaking
existing software into a platform for building new software from
scratch. This would likely end up resembling existing tools for building
spreadsheet-driven applications (discussed more below), but with an
additional focus on customization by end users of the software.

\hypertarget{web-customization}{%
\subsection{Web customization}\label{web-customization}}

Wildcard's goals are closely shared with other systems that provide
interfaces in the browser for end users to customize websites.

\hypertarget{structured-augmentation}{%
\subsubsection{Structured augmentation}\label{structured-augmentation}}

Wildcard's solution approach is most similar to other tools that
identify structured data in a web page and then use that structure to
support end user customization of the page.

Sifter \autocite{huynh2006} enables users to sort and filter lists of
data on web pages, which resembles Wildcard's sorting functionality. The
underlying mechanism is similar, since Sifter extracts structured data
from the page to enable its user-facing functionality. However, the
systems also have significant differences:

\begin{itemize}
\tightlist
\item
  Wildcard hides data extraction from end users by having programmers
  develop site adapters, whereas Sifter uses a combination of automated
  heuristics and interactive user feedback to involve end users in
  extracting data. User studies of Sifter indicated that users were
  surprised by the need to participate in the data augmentation process,
  which influenced our decision to hide this process from the user. We
  suspect Wildcard's workflow results in a simpler end user experience,
  but at the cost of only working with a smaller subset of supported
  websites.
\item
  Wildcard supports a broad set of customizations. Sifter only supports
  sorting and filtering.
\item
  Wildcard shows the structured data table directly to the user. Sifter
  only shows sorting controls, and does not reveal the underlying data
  table.
\end{itemize}

Thresher \autocite{hogue2005} enables users to create wrappers which map
unstructured website content to Semantic Web content. Like Wildcard and
Sifter, Thresher augments the page: once semantic content has been
identified, it creates context menus in the original website which allow
users to take actions based on that content. Wildcard and Thresher focus
on complementary parts of the customization process. Thresher aims to
enable end users to create content wrappers, but the actions available
on the structured data are determined in advance. Conversely, Wildcard
delegates wrapper creation to programmers, but gives end users more
flexibility to use the structured data in an open-ended way.

\hypertarget{sloppy-augmentation}{%
\subsubsection{Sloppy augmentation}\label{sloppy-augmentation}}

``Sloppy programming'' \autocite{little2010} tools like Chickenfoot
\autocite{bolin2005} and Coscripter \autocite{leshed2008} enable users
to create web automation scripts without directly interacting with the
DOM. The scripts can perform actions like filling in text boxes and
clicking on buttons. Users express the desired page elements in natural,
informal terms (e.g.~writing ``the username box'' to represent the
textbox closest to the label ``username''), and then the system uses a
set of heuristics to determine which elements most likely match the
user's intent. This approach allows for expressing a wide variety of
commands with minimal training, but it also has downsides
\autocite{little2010}:

\begin{itemize}
\tightlist
\item
  \emph{Reliability}: It can be difficult to know whether a command will
  consistently work over time. In addition to changes to the website,
  changes to the system's heuristics can also cause unexpected changes
  in behavior.
\item
  \emph{Discoverability}: it can be hard for users to find discover
  possible commands.
\end{itemize}

Wildcard offers a sharp contrast to sloppy programming, instead choosing
to expose a high degree of structure through the spreadsheet table.
Wildcard offers more consistency; for example, clicking a sort header
will always work correctly as long as the site adapter is maintained.
Wildcard also offers clearer affordances for what types of actions are
possible. On the other hand, Wildcard's explicit site adapter approach
means that fewer websites can be customized, and also that users can't
perform customizations if the relevant data is not exposed in the
spreadsheet.

Another difference between Wildcard and these tools is that there is no
way for users to express imperative workflows with sequences of actions
in Wildcard. Chickenfoot and Coscripter have shown these types of
workflows to be useful in practice so it might be worth considering how
to support them, but it's not clear how such workflows could fit into
Wildcard. Spreadsheets have a fundamentally different computation model,
and site adapters created by programmers cannot easily account for every
possible action a user might want to perform in a page.

\hypertarget{spreadsheet-driven-applications}{%
\subsection{Spreadsheet-driven
applications}\label{spreadsheet-driven-applications}}

Many researchers and companies have explored the idea of
\emph{spreadsheet-driven applications}: using spreadsheets as a backing
data store and computation layer for interactive web applications,
enabling end users to create such applications more easily. Research
projects like Object Spreadsheets \autocite{mccutchen2016}, Quilt
\autocite{benson2014}, Gneiss \autocite{chang2014}, and Marmite
\autocite{wong2007}, as well as commercial tools like Airtable
\autocite{zotero-79} and Glide \autocite{zotero-81} allow users to view
data in a spreadsheet table, compute over the data using formulas, and
then connect the table to a GUI. Because many users are already familiar
with using spreadsheets, this way of creating applications tends to be
much easier than traditional software methods; for example, in a user
study of Quilt, many users were able to create applications in under 10
minutes, even if they expected it would take them many hours.

Wildcard builds on this idea, but applies it to modifying existing
applications, rather than building new applications from scratch. For
many people, we suspect that tweaking existing applications provides
more motivation as a starting point for programming than creating a new
application from scratch.

An important design decision for tools in this space is how far to
deviate from traditional spreadsheets like Microsoft Excel or Google
Sheets. Quilt and Glide use existing spreadsheet software as a backend,
providing maximum familiarity for users, and even compatibility with
existing spreadsheets. Gneiss has its own spreadsheet implementation
with additional features useful for building GUIs. Marmite provides a
live data view that resembles a spreadsheet, but programming is actually
done using a separate data flow pane rather than spreadsheet formulas.
Marmite's approach led to some confusion in a user study, because users
expected behavior more similar to spreadsheets \autocite{wong2007}.
Airtable deviates the furthest: although the user interface resembles a
spreadsheet, the underlying structure is a relational database with
typed columns. Wildcard's table is most similar to Airtable; the
structure of a relational table is most appropriate for most data in
websites, and we have not yet found a need for arbitrary untyped cells.

\hypertarget{web-scraping-data-extraction}{%
\subsection{Web scraping / data
extraction}\label{web-scraping-data-extraction}}

Web scraping tools focus on extracting structured data out of
unstructured web pages. Web scraping is closely related to the
implementation of Wildcard, but has different end goals: web scraping
generally extracts static data for processing in another environment,
whereas Wildcard customizes the application's UI by maintaining a
bidirectional connection between the extracted data and the page.

Web scraping tools differ in how much structure they attempt to map onto
the data. Some tools like Rousillon \autocite{chasins2018} extract data
in a minimally structured relational format; other tools like Piggy Bank
\autocite{huynh2005} more ambitiously map the data to a rich semantic
schema. In Wildcard, we chose to avoid semantic schemas, in order to
minimize the work associated with creating a site adapter.

In the future, we might try to integrate existing web scraping tools, to
help create more reliable site adapters for Wildcard with less work, and
to open up adapter creation to end users. Although Wildcard has
additional requirements beyond traditional web scraping as discussed in
Section~\ref{sec:implementation}, there is still enough overlap that
existing tools might prove helpful.

\hypertarget{sec:conclusion}{%
\section{Conclusion and Future Work}\label{sec:conclusion}}

In this paper, we have presented \emph{spreadsheet-driven
customization}, a technique that enables end users to customize software
by augmenting an application's UI with a spreadsheet. We still have many
open questions which we hope to answer through targeted development and
usage of Wildcard.{\footnote{Up to date information about Wildcard is
  available on the project website
  (\url{https://geoffreylitt.com/wildcard}), which includes a signup for
  receiving updates and notifications about a public release. We also
  welcome hearing from readers who might be interested in being private
  beta testers or who have feedback on the project.}}

The most important question is: what are the limits of this
computational model? What types of useful customizations can it support
or not support? While initial demos suggest a variety of use cases, we
plan to develop more site adapters and demos to explore this question
further. We will start by privately testing the system with our own
needs and then eventually deploy the tool publicly once it has a stable
API and can support a critical mass of sites and use cases. We also plan
to run usability studies to evaluate and improve the design of the tool.

We suspect that some areas of the current model may prove overly
simplistic. For example, Wildcard's data model currently shows a single
table at a time, without any notion of relationships between tables. A
richer data model with foreign keys might help support certain use
cases. For designing a spreadsheet interface on top of a richer
relational model, we could learn from other systems with this design
\autocite{mccutchen2016,bakke2016}.

Another question is how easily site adapters can be created and
maintained for real websites, which often include complex markup and
change frequently. We plan to explore this question by creating adapters
for sites with different data structures and in different domains, and
perhaps running user tests with programmers. Possible future
improvements we've considered include developing automated heuristics to
assist with the adapter creation process and developing new abstractions
that make it easier for programmers to efficiently create robust
adapters.

We hope that spreadsheet-driven customization can help the Web reach its
full potential as a dynamic medium that users can mold to their own
needs.

\acks
Thanks to Glen Chiacchieri, David Karger, Steve Krouse, Rob Miller, Arvind Satyanarayan, Daniel Windham, and Maggie Yellen for providing valuable feedback on this work.

\printbibliography

\end{document}

% Local Variables:
% TeX-engine: luatex
% End: